---
title: "Exploratory Data Analysis on SF Crime Data"
author: "M. Demas"
date: "15 February 2016"
output: pdf_document
---

# Days of the Week
```{r relevel-day-factor}
train$DayOfWeek <- as.character(train$DayOfWeek)
train$DayOfWeek <- factor(train$DayOfWeek, levels=c("Sunday", "Monday", 
                                                    "Tuesday", "Wednesday", 
                                                    "Thursday", "Friday", 
                                                    "Saturday"))
```

```{r summarize-data-into-count}
library(dplyr)
df <- train %>% 
  group_by(PdDistrict, DayOfWeek, Category) %>% 
  summarise(Count = n())
```

Crimes of day of the week is not very enlightening. So I conditioned it by District.
```{r plot-conditioned-on-district}
fig_dir = '~/Box Sync/KaggleSFCrime/Figures/'
setwd(fig_dir)

library(ggplot2)
ggplot(df, aes(x=DayOfWeek, y=Count)) + geom_bar(stat="identity")
ggplot(df, aes(x=DayOfWeek, y=Count)) + 
  geom_bar(stat="identity") + 
  facet_wrap(~PdDistrict) +
  ggtitle("Count of Crimes for Each Day of the Week Conditioned on District")
ggsave('CrimeCountPerDayOfWeekInEachDistrict_2016-02-16.png')
```

It would be interesting to understand how well SF is "mapped" by the crimes. A first scatter plot reveals some outliers Southern with coordinates long = -120.5, lat = 90 (perhaps they represent crimes across the bay?). I filtered out those values, and plotted again. The level of coverage seems fairly comprehensive. However, there appear to be some District misclassifications.
```{r scatter-plot-coordinates}

ggplot(train, aes(x=X, y=Y, color=PdDistrict)) + 
  geom_point(position='jitter') +
  ggtitle("Scatter Plot of Crimes (long.,lat.) Conditioned on District")
ggsave('CrimeScatterPlotInEachDistrict_2016-02-16.png')

ggplot(train %>% filter(X < -122.0, Y < 50), aes(x=X, y=Y, color=PdDistrict)) + 
  geom_point(position='jitter') +
  ggtitle("Scatter Plot of Crimes (long.,lat.) Conditioned on District")
ggsave('CrimeScatterPlotInEachDistrict_2016-02-16.png')
